\begin{abstract}
Could technology be responsible for part of the rise in income inequality over the past 30 years? This research is motivated by the fact that, while technology can make workers more productive, it can also render other workers unneccessary by automating some tasks. By changing demand patterns for different kinds of labor, it is possible that technology is responsible for an increase in the demand and wage of certain types of work, and a reduction in the wage and number of others. Research from the United States and Europe suggests that technological change has indeed caused such a `polarization' of the wage distribution.

In this thesis, we assess the evidence for polarization in Australia. We first consider the standard model of skill-biased technical change, and show that it only poorly fits the observed data. We link the wage share of middle-skilled occupations to investment in electronic and electrical capital goods, and also demonstrate a relationship between qualitative properties of certain jobs, and changes in the wage distribution. We find that jobs of the kind most likely to be impacted by technology, so-called `routine' jobs, have suffered the greatest decline in income over the past 30 years, and that the penalty associated with routine work is concentrated in upper wage quantiles. Finally, we find a large change in the value of occupations involving managerial tasks, an effect that may or may not be associated with technical change.
\end{abstract}
