\title{Technical Change \& Wages in Australia}
\author{Alex Cooper}
\date{\parbox{\linewidth}{\centering%
  November, 2013\endgraf\bigskip\vspace{5cm}
  Submitted as a partial requirement for the completion of the degree of Bachelor~of~Economics~(Hons).
}}
\maketitle
\vspace{30pt}

%\frontmatter
\clearpage
\thispagestyle{empty}
\vspace*{\fill}
\begin{center}
\begin{minipage}{.8\textwidth}
%\begin{quotation}
        \begin{center}
          { \Large \bf Acknowledgments }
        \end{center}
I am deeply grateful to the many academics and friends who have led me through this project.  To Dr Michael Dobbie, Dr Chris Heaton, and Dr Roselyne Joyeux: thank you so much for your unbounded patience, sage advice, and for giving so generously of your time. Without your assistance, this project would never have been completed. Thank you in particular for curbing my optimism when that was called for, and for prodding me to persist when things weren't going as I'd planned. 

I'm also grateful to Dr Stephane Mahuteau for advice about data sources, and Dr Chris Bidner for discussing an early version of this work at the National Honours Colloquium.

To Dr Tony Bryant, thank you for the endless advice and moral support you offered throughout the year, and for listening for my tales of woe about statistical software. Thanks especially for helping me to see the lighter side of the many challenges that sprang up, and encouraging me to laugh off the stress.

And finally, to Kelsey, whose patience I tested in every conceivable way as I wrote this paper. I would not have made it without your endless support.

%\end{quotation}
\end{minipage}
\end{center}
\vfill % equivalent to \vspace{\fill}
\clearpage


%%% Local Variables: 
%%% mode: latex
%%% TeX-master: "paper"
%%% End: 
